\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-08-09 paper.tex
Partial proofreading, light reogranization of chap 3

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}

@@

SPARK is an \emph{amodal} geometric model of the environment that serves both
as basis for the fusion of the perception modalities and as bridge with the
symbolic layer. This geometric model is continuously updated at run-time by the
robot based on its sensors (in our experiments, objects are identified and
localised through 2D barcodes, while humans are tracked with Kinect-like
devices, optionally assisted by motion capture to accurately track the head
motion, which is required to compute what the human is looking at).

@@

\paragraph{Symbolic facts production:}

@@

Reasoning about human perspective allow to compute facts such as:
\stmt{GREY\_TAPE isBehind HUMAN1}, \stmt{GREY\_TAPE isVisibleBy
  HUMAN1}.

Figure~\ref{fig::reach-ex} illustrates different situations for the
\textit{reachable} relation. In this case, the robot and its human partner
are placed face to face, in a table-top setup (Figure~\ref{fig::reach-ex}.1).
The robot first estimates if the small grey box is reachable to itself. This is
done by finding a collision free posture to reach the object
(Figure~\ref{fig::reach-ex}.2). Next the robot switches to the human's
perspective to estimate if the same object is reachable to the human as well.
In the last scene, the human moves towards his left, farther from the object
(Figure~\ref{fig::reach-ex}.4). The situation is then reevaluated. In this
occasion though, the reasoner cannot find a satisfactory posture for the human
to reach the box because he is too far from the target.

\begin{figure*}[!t]
	\centering
	\includegraphics[width=1.0\textwidth]{figs/reachex.jpg}	
	\caption{An example illustrating the \textit{reachable} relation. The relation is computed from the perspectives of both the robot and the human. The computed posture at each step is illustrated with a global view of the scene (top), and from a closest view (bottom).}
\label{fig::reach-ex}
\end{figure*}

\begin{figure*}[ht!]
   \label{fig:sparkSubfigures}
   \begin{center}
%
       \subfigure[Initial state]{%
%           \label{fig:wuweiPhoto}
           \includegraphics[width=0.5\textwidth]{./figs/etat2-P1010769_brightened-v2.jpg}
       }%
       \subfigure[3d model view of initial state]{%
%          \label{fig:sparkScreenshot}
          \includegraphics[width=0.43\textwidth]{./figs/etat2_photo.png}
       }\\ %  ------- End of the first row ----------------------%
%
   \end{center}

   \caption{%
     In this situation, there are three tapes on the table. Two tapes
     are only reachable by the robot: the LOTR\_TAPE (black in the 3d
     model) and GREY\_TAPE. The third tape WALLE\_TAPE (white in the
     3d model) and the trashbin PINK\_TRASHBIN are only reachable by
     the human HUMAN1. All tapes are on the table TABLE.  }%

@@

\paragraph{Hypotheses on objects states and positions}
It is sometimes difficult or even impossible to see and/or track an
object in certain states. This happens, for instance, when the object
has been put in a container, when it is in the robot gripper or in the
human hand, and more generally in any state in which it is hidden by
something else. Our robot has a model of the possible symbolic states
for an object (whether the object is on a furniture, in an agent hand,
in a container, etc.).  According to the robot perception of what has
happened since the object was last seen, the robot tries to maintain a
belief of the current possible symbolic states and their associated
probabilities for this object. Such information can be used to update
the beliefs using input from exploration, dialog, human visual
focus,\ldots
SPARK currently provides a simple implementation of such a functionality. The
only managed hypotheses are \emph{in container} and \emph{in agent hand}. We
can have only one hypothesis at the same time. Hypothesis validity is checked
geometrically in case of incoming perception values.

\paragraph{Primitive action recognition}
Monitoring human activity is crucial to maintain a coherent state of
the world. Full human action and activity monitoring is a difficult
task that requires knowledge and reasoning both on high level facts
like goals, intentions and plans, as well as bottom-up data from agent
and object motions. Simple temporal and geometric reasoning on human
hand trajectories and potential objects placements can provide some
useful clues for high level human monitoring processes. We call this
temporal and geometric reasoning \emph{primitive action recognition}.

For example, a \emph{pick}, a \emph{throw} or a \emph{place} action
can be recognized by observing that an object on table and an empty
human hand are close to each other, or that the human hand holding an
object is close to a container, etc. Human hand position is either
directly perceived or inferred from its initial perceived trajectory.
We have a simple implementation of such a primitive action recognition
in SPARK that relies on monitoring human hand and its motion near
objects or above containers.

@@

SPARK enables \emph{perspective-taking}: spatial relations between entities can
be computed from different viewpoints, which let the robot build for each
agents it interacts with a different, \emph{perspective-aware} symbolic model
of the environment. These models are separately stored in the knowledge base.

@@

As a result, the robot stores models of the environment either in the
\emph{ego-centric} reference frame (from the robot perspective) or in the
\emph{allo-centric} frame (addressee-centred).

\subsubsection{Symbolic facts and beliefs management}

The facts produced by the geometric and temporal reasoning component
are stored in a central symbolic knowledge base, called ORO. Besides
acting as a facts database, the ORO platform~\cite{Lemaignan2010}
exposes several functions: operations on knowledge statements relying
on inference (through a continuous first-order logic classification
process), management of \emph{per-agent} symbolic models, and also
higher cognitive and human-robot interaction related functionalities
like categorization of sets of concepts, profiles of memory (that
enable the robot to ``forget'' about some facts), natural language
grounding~\cite{Lemaignan2011}\ldots.

@@

ORO also provides an event mechanism that allows components to be
triggered when specific events occur. A component can
for instance subscribe to events of kind \stmt{?agent isVisible
  true, ?agent type Human}. As soon as the perception layer detects a
human in the robot's field of view and accordingly updates the
knowledge base, the executive layer would be triggered back. The
event framework also takes advantage of the inference capabilities of
ORO. Thus an event can be indirectly triggered if its triggering
conditions can be inferred to be true.

\subsubsection{Context Management Policies}

\fxfatal{Rewrite this section to level up AI aspects (Mathieu)}

Building, updating and maintaining a correct state of the world at
geometric and symbolic level is crucial to the capacity of the robot
to carry on successfully a multi-step interaction with a human. Tight
integration between the robot controller and the geometric and
temporal reasoning functions in SPARK and symbolic facts and beliefs
management in ORO is central.

Robot reasons on possible positions for non perceived
objects. These hypotheses could be updated using new input from dialogue, human
action and focus of attention. Currently, we manage at most one hypothesis per
object. This hypothesis is produced by robot controller through an inference on
robot or human action. In case of perception conflicts with low probability for
the current hypothesis, robot controller will break this hypothesis and delete
corresponding symbolic fact in the ontology.

@@

Natural language input (in experiments, we rely on an Android-based interface,
with Google speech recognition) is parsed into a grammatical structure, and
atoms of each sentence are resolved with the help of the ontology to ground
concepts like objects (\ie when a user says ``pick the can'', resolve to which
instance of \emph{can} the user is referring to) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.

The system supports quantification (``give me \{a | the | some | all | any |
...\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal}), interprets common temporal and
place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a
certain extend \emph{states} (``I'm tired'') into \emph{experiences}
(\stmt{HUMAN experiences state\_1, state\_1 hasFeature tired}).

@@

First, particular steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when a set of instances
matches a category (the human says ``give me the bottle'' and the robot knows
about three bottles), the module may decide (it actually depends on the
quantifier preceding the class) to discard some of them based on their
\emph{visibility} for the speaker (implicit communication context built on the
human posture).

@@

Note that, while the system benefits from the complementary modalities, they
are not all required. The dialogue system would for instance happily run with
only the verbal modality, at the cost of weaker interaction.

@@

Our execution controllers rely on symbolic task planning to convert faraway
desires into a succession of atomic actions. We use in our architecture the
HATP planner (for Human Aware Task Planner~\cite{Alili2009}).  HATP is based on
a Hierarchical Task Network (HTN) refinement, which performs an iterative task
decomposition into sub-tasks until reaching atomic actions. The \emph{planning
domain} defines the set of methods describing how to decompose a task and
represents the procedural knowledge of the robot.

In order to devise how a given goal can be accomplished, the robot has
to elaborate a plan,~\textit{i.e.} a set of actions to be achieved by
the robot and its human partners.  This is the role of a HATP
\cite{Alili2008} (for Human Aware Task Planner).  HATP is based on a
Hierarchical Task Network (HTN) refinement which performs an iterative
task decomposition into sub-tasks until reaching atomic
actions~\cite{Nau2003}.  The planning domain defines a set of methods
describing how to decompose a task and can be seen as the Howto
knowledge of the robot.  HATP is able to produce plans for the robot's
actions as well as for the other participants (humans or robots). It
can be tuned by setting up different costs depending on the actions to
apply and by taking into account a set of constraints called social
rules. This tuning aims at adapting the robot's behavior according to
the desired level of cooperation of the robot.

@@

>>>>

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\al}{{\textit{et al.\ }}}

@@

SPARK is an \emph{amodal}~\cite{Mavridis2006} geometric model of the
environment that serves both as basis for the fusion of the perception
modalities and as bridge with the symbolic layer. This geometric model is
continuously updated at run-time by the robot based on its sensors (in our
experiments, objects are identified and localised through 2D barcodes, while
humans are tracked with Kinect-like devices, optionally assisted by motion
capture to accurately track the head motion, which is required to compute what
the human is looking at).

\fxnote{Add 2-3 lines to summarize the AI highlights of SPARK}

@@

\paragraph{Symbolic facts production}

@@

Reasoning about human perspective allow to compute facts such as:
\stmt{GREY\_TAPE isBehind HUMAN1}, \stmt{GREY\_TAPE leftOf
  HUMAN1}.

Figure~\ref{fig::reach-ex} illustrates different situations for the
\textit{reachable} relation. The robot and its human partner are placed face to
face (Figure~\ref{fig::reach-ex}.1).  The robot first estimates if the small
grey box is reachable to itself. This is done by finding a collision free
posture to reach the object (Figure~\ref{fig::reach-ex}.2). Next the robot
switches to the human's perspective to estimate if the same object is reachable
to the human as well.  In the last scene, the human moves towards his left,
farther from the object (Figure~\ref{fig::reach-ex}.4). The situation is then
reevaluated. In this occasion though, the reasoner cannot find a satisfactory
posture for the human to reach the box because he is too far from the target.

\begin{figure*}[!t]
	\centering
    \subfigure[] { \includegraphics[width=0.2\textwidth]{figs/reachex-1.jpg} }
    \subfigure[] {\includegraphics[width=0.2\textwidth]{figs/reachex-2.jpg} }
    \subfigure[] {\includegraphics[width=0.2\textwidth]{figs/reachex-3.jpg} }
    \subfigure[] {\includegraphics[width=0.2\textwidth]{figs/reachex-4.jpg} }
	\caption{An example illustrating the \textit{reachable} relation. The relation is computed from the perspectives of both the robot and the human. The computed posture at each step is illustrated with a global view of the scene (top), and from a closest view (bottom).}
\label{fig::reach-ex}
\end{figure*}

\begin{figure*}[ht!]
   \begin{center}
%
       \subfigure[Initial state]{%
%           \label{fig:wuweiPhoto}
           \includegraphics[width=0.5\textwidth]{./figs/etat2-P1010769_brightened-v2.jpg}
       }%
       \subfigure[3d model view of initial state]{%
%          \label{fig:sparkScreenshot}
          \includegraphics[width=0.43\textwidth]{./figs/etat2_photo.png}
       }\\ %  ------- End of the first row ----------------------%
%
   \end{center}

   \caption{%
     In this situation, there are three tapes on the table. Two tapes
     are only reachable by the robot: the LOTR\_TAPE (black in the 3d
     model) and GREY\_TAPE. The third tape WALLE\_TAPE (white in the
     3d model) and the trashbin PINK\_TRASHBIN are only reachable by
     the human HUMAN1. All tapes are on the table TABLE.  }%
   \label{fig:sparkSubfigures}

@@

\paragraph{Hypotheses on objects states and positions}
It is sometimes difficult or even impossible to see and/or track an
object in certain states. This happens, for instance, when the object
has been put in a container, when it is in the robot gripper or in the
human hand, and more generally in any state in which it is hidden by
something else. Our robot has a model of the possible symbolic states
for an object (whether the object is on a furniture, in an agent hand,
in a container, etc.).  According to the robot perception of what has
happened since the object was last seen, the robot tries to maintain a
belief of the current possible symbolic states and their associated
probabilities for this object. Such information can be used to update
the beliefs using input from exploration, dialog, human visual
focus,\ldots \fxnote{Paragraph need write. Tell what the module actually does, not what it could do}

SPARK currently provides a simple implementation of such a functionality. The
only managed hypotheses are \emph{in container} and \emph{in agent hand}. We
can have only one hypothesis at the same time. Hypothesis validity is checked
geometrically in case of incoming perception values.

\paragraph{Primitive action recognition} Monitoring human activity is crucial
to maintain a coherent state of the world. Full human action and activity
monitoring is a task that requires knowledge and reasoning both on high level
facts like goals, intentions and plans, as well as bottom-up data from agent
and object motions. Simple temporal and geometric reasoning on human hand
trajectories and potential objects placements can provide some useful clues for
high level human monitoring processes. We call this temporal and geometric
reasoning \emph{primitive action recognition}.

For example, a \emph{pick}, a \emph{throw} or a \emph{place} action
can be recognized by observing that an object on table and an empty
human hand are close to each other, or that the human hand holding an
object is close to a container, etc. Human hand position is either
directly perceived or inferred from its initial perceived trajectory\fxwarning{we really do that?}.
We have a simple implementation of such a primitive action recognition
in SPARK that relies on monitoring human hand and its motion near
objects or above containers.

@@

SPARK enables \emph{perspective-taking}: spatial relations between entities can
be computed from different viewpoints, which let the robot build for each
agents it interacts with a different, \emph{perspective-aware} symbolic model
of the environment. These models are separately stored in the knowledge base.
\fxwarning{Rephrase: already mentioned before}

@@

As a result, the robot stores models of the environment either in the
\emph{ego-centric} reference frame (from the robot perspective) or in 
\emph{allo-centric} frames (addressee-centred).

\subsubsection{Symbolic facts and beliefs management}
\fxwarning{Do we need this section here?}

The facts produced by the geometric and temporal reasoning component
are stored in the central symbolic knowledge base.

@@

@@

Natural language input (in experiments, we rely on an Android-based interface,
with Google speech recognition) is parsed into a grammatical structure, and
atoms of each sentence are resolved with the help of the ontology to ground
concepts like objects (\ie when a user says ``pick the can'', resolve to which
instance of \emph{can} the user is referring to) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.\fxwarning{Severin: Develop this section}

The system supports quantification (``give me \{a | the | some | all | any |
...\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal} and added to th eknowledge pool),
interprets common temporal and place adverbs (like \emph{above} or
\emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm
tired'') into \emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1
hasFeature tired}).

@@

First, particular steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when several instances
matches a category (the human says ``give me the bottle'' and the robot knows
about three bottles), the module may decide (it actually depends on the
quantifier preceding the class) to discard some of them based on their
\emph{visibility} for the speaker (implicit communication context built on the
human posture).

@@

Note that, while the system benefits from the complementary modalities, they
are not all required. The dialogue system can run with only the verbal
modality, at the cost of a simpler interaction.

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the
HATP planner (for Human Aware Task Planner~\cite{Alili2008, Alili2009}).  HATP
is based on a Hierarchical Task Network (HTN) refinement, which performs an
iterative task decomposition into sub-tasks until reaching atomic
actions\fxwarning{Rephrase: the definition is actually generic to HTN, not HATP
specfic}. The \emph{planning domain} defines the set of methods describing how
to decompose a task and represents the procedural knowledge of the robot. It is
stored outside of the central knowledge base.

HATP is able to produce plans for the robot's actions as well as for the other
participants (humans or robots). It can be tuned by setting up different costs
depending on the actions to apply and by taking into account a set of
constraints called social rules. This tuning aims at adapting the robot's
behavior according to the desired level of cooperation of the robot.

@@

\paragraph{Context Management Policies} \fxfatal{Rewrite this section to level up AI aspects (Mathieu)}
Building, updating and maintaining a correct state of the world at
geometric and symbolic level is crucial to the capacity of the robot
to carry on successfully a multi-step interaction with a human. Tight
integration between the robot controller and the geometric and
temporal reasoning functions in SPARK and symbolic facts and beliefs
management in ORO is central.

Robot reasons on possible positions for non perceived objects. These hypotheses
could be updated\fxwarning{are they?} using new input from dialogue, human action and focus of
attention. Currently, we manage at most one hypothesis per object. This
hypothesis is produced by robot controller through an inference on robot or
human action. In case of perception conflicts with low probability for the
current hypothesis, robot controller will break this hypothesis and delete
corresponding symbolic fact in the ontology.
